---
title: "Application Exponential Distribution"
output: html_notebook
---
 
 

```{r error=FALSE , warning=FALSE,echo=FALSE}
 library(tidyverse)  
 library(lubridate)
 library(plotly)
 library(hrbrthemes)
```

 
# Load Data 
```{r error=FALSE , warning=FALSE,echo=FALSE}
results <- read.csv("results.csv")
str(results)
```
# Univariae Visualization

## Outliers Proportion
```{r}
p1 <- results %>%   ggplot(aes(y = (outliersPro ))) +
  geom_boxplot() + labs(title = "Outliers Proportion")

p2 <- results %>%   ggplot(aes(x = (outliersPro ))) +
  geom_histogram() + labs(title = "Outliers Proportion")

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )

```


## Estimated Rate Before Handling
```{r}
p1 <- results %>%   ggplot(aes(y = (estimated_rate_Before_Handling ))) +
  geom_boxplot() + labs(title = "Estimated Rate Before Handling")

p2 <- results %>%   ggplot(aes(x = (estimated_rate_Before_Handling ))) +
  geom_histogram() + labs(title = "Estimated Rate Before Handling" , subtitle = "scale X >> Log(10) "           ) + scale_x_log10()

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )

```


## Estimated Rate After Handling {Mean}
```{r}
p1 <- results %>%   ggplot(aes(y = (estimated_rate_After_mean      ))) +
  geom_boxplot() + labs(title = "Estimated Rate After Handling {Mean}")

p2 <- results %>%   ggplot(aes(x = (estimated_rate_After_mean      ))) +
  geom_histogram()  + labs(title = "Estimated Rate After Handling {Mean}"
                          , subtitle = "scale X >> Log(10) "
                           ) + scale_x_log10()
  

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )

```



```{r}
results %>%   select(estimated_rate_Before_Handling , 
                     estimated_rate_After_Q_b_F_C , 
                     estimated_rate_After_mean , 
                     estimated_rate_After_median 
                     ) %>% 
  gather("Method" , "estimatedRate" , estimated_rate_Before_Handling , estimated_rate_After_Q_b_F_C , estimated_rate_After_mean ,estimated_rate_After_median  ) %>% 
  ggplot(aes(x = (estimatedRate))) + 
     geom_histogram( aes( fill  = as.factor(Method))) + 
   labs(title = "Estimated Rate After Handling "
                          , subtitle = "scale X >> Log(10) " ) + scale_x_log10()+
  

  facet_wrap(.~Method)


```



# Relation Between Outliers Proportion and P-value Before 
```{r}

p1 = results %>%
  mutate(fitted = ifelse(Before_P_value > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x = results$outliersPro , y= Before_P_value)) + 
  geom_point(aes(color = fitted , alpha = .9)) +
    labs(title = "Outliers Proportion And P_value Before"
                             ) 

p2 = results %>%
  mutate(fitted = ifelse(Before_P_value > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x= fitted)) + 
  geom_bar(aes(fill=fitted)) +
    labs(title = "Outliers Proportion And P_value Before"
                             )

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )
```


# Relation Between Outliers Proportion and P-value After {Mean} 
```{r}

p1 = results %>%
  mutate(fitted = ifelse(mean_P_value > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x = results$outliersPro , y= results$mean_P_value)) + 
  geom_point(aes(color = fitted , alpha = .5) ) +
    labs(title = "Outliers Proportion And P_value After"
                             )  

p2 = results %>%
  mutate(fitted = ifelse(mean_P_value > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x= fitted)) + 
  geom_bar(aes(fill=fitted)) +
    labs(title = "Outliers Proportion And P_value After"
                             )

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )

```




# Relation Between Outliers Proportion and P-value After {Median} 
```{r}

p1 = results %>%
  mutate(fitted = ifelse(median_P_value   > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x = results$outliersPro , y= results$median_P_value)) + 
  geom_point(aes(color = fitted , alpha = .5) ) +
    labs(title = "Outliers Proportion And P_value After"
                             )  

p2 = results %>%
  mutate(fitted = ifelse(median_P_value > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x= fitted)) + 
  geom_bar(aes(fill=fitted)) +
    labs(title = "Outliers Proportion And P_value After"
                             )

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )

```




# Relation Between Outliers Proportion and P-value After {Q_b_F_C_P} 
```{r}

p1 = results %>%
  mutate(fitted = ifelse(Q_b_F_C_P_value   > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x = results$outliersPro , y= results$Q_b_F_C_P_value)) + 
  geom_point(aes(color = fitted , alpha = .5) ) +
    labs(title = "Outliers Proportion And P_value After"
                             )  

p2 = results %>%
  mutate(fitted = ifelse(Q_b_F_C_P_value > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x= fitted)) + 
  geom_bar(aes(fill=fitted)) +
    labs(title = "Outliers Proportion And P_value After"
                             )

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )

```




 


