---
title: "Application Exponential Distribution"
output:
  html_notebook: default
  pdf_document: default
---
 
 

```{r error=FALSE , warning=FALSE,echo=FALSE}
 library(tidyverse)  
 library(lubridate)
 library(plotly)
 library(hrbrthemes)
```

 
# Load Data 
```{r error=FALSE , warning=FALSE,echo=FALSE}
results <- read.csv("results.csv")
str(results)
```
# Univariae Visualization

## Outliers Proportion
```{r}
p1 <- results %>%   ggplot(aes(y = (outliersPro ))) +
  geom_boxplot() + labs(title = "Outliers Proportion")

p2 <- results %>%   ggplot(aes(x = (outliersPro ))) +
  geom_histogram(fill = "steelblue") + labs(title = "Outliers Proportion")+
  theme_minimal()

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )

```


## Estimated Rate Before Handling
```{r}
p1 <- results %>%   ggplot(aes(y = (estimated_rate_Before_Handling ))) +
  geom_boxplot() + labs(title = "Estimated Rate Before Handling")

p2 <- results %>%   ggplot(aes(x = (estimated_rate_Before_Handling ))) +
  geom_histogram() + labs(title = "Estimated Rate Before Handling" , subtitle = "scale X >> Log(10) "           ) + scale_x_log10()

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )

```

```{r}
qqnorm((results$estimated_rate_Before_Handling), pch = 1, frame = FALSE)
qqline((results$estimated_rate_Before_Handling), col = "steelblue", lwd = 2)

```



## Estimated Rate After Handling {Mean}
```{r}
p1 <- results %>%   ggplot(aes(y = (estimated_rate_After_mean      ))) +
  geom_boxplot() + labs(title = "Estimated Rate After Handling {Mean}")

p2 <- results %>%   ggplot(aes(x = (estimated_rate_After_mean      ))) +
  geom_histogram()  + labs(title = "Estimated Rate After Handling {Mean}"
                          , subtitle = "scale X >> Log(10) "
                           ) + scale_x_log10()
  

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )

-```



```{r}
results %>%   select(estimated_rate_Before_Handling , 
                     estimated_rate_After_Q_b_F_C , 
                     estimated_rate_After_mean , 
                     estimated_rate_After_median 
                     ) %>% 
  gather("Method" , "estimatedRate" , estimated_rate_Before_Handling , estimated_rate_After_Q_b_F_C , estimated_rate_After_mean ,estimated_rate_After_median  ) %>% 
  ggplot(aes(x = (estimatedRate))) + 
     geom_histogram( aes( fill  = as.factor(Method))) + 
   labs(title = "Estimated Rate After Handling "
                          , subtitle = "scale X >> Log(10) " ) + scale_x_log10()+
  

  facet_wrap(.~Method)


```




# Relation Between Outliers Proportion and P-value Before 
```{r}

p1 = results %>%
  mutate(fitted = ifelse(Before_P_value > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x = results$outliersPro , y= Before_P_value)) + 
  geom_point(aes(color = fitted , alpha = .9)) +
    labs(title = "Outliers Proportion And P_value Before"
                             ) 

p2 = results %>%
  mutate(fitted = ifelse(Before_P_value > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x= fitted)) + 
  geom_bar(aes(fill=fitted)) +
    labs(title = "Outliers Proportion And P_value Before"
                             )

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )
```
```{r}

  

 
```


# Relation Between Outliers Proportion and P-value After {Mean} 
```{r}

p1 = results %>%
  mutate(fitted = ifelse(mean_P_value > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x = results$outliersPro , y= results$mean_P_value)) + 
  geom_point(aes(color = fitted , alpha = .5) ) +
    labs(title = "Outliers Proportion And P_value After"
                             )  

p2 = results %>%
  mutate(fitted = ifelse(mean_P_value > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x= fitted)) + 
  geom_bar(aes(fill=fitted)) +
    labs(title = "Outliers Proportion And P_value After"
                             )

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )

```


Parametric Test for Association Between { Fitted Value Before And Fitted Values After {Mean} }
```{r}

ddd <-results %>%
  mutate(fittedBefore = ifelse(Before_P_value > 0.05,TRUE,FALSE)) %>%
  mutate(fittedAfter = ifelse(mean_P_value > 0.05,TRUE,FALSE)) 

 
library(MASS)
tb1<-table(ddd$fittedBefore, ddd$fittedAfter)
tb1
chisq.test(tb1)

```



# Relation Between Outliers Proportion and P-value After {Median} 
```{r}

p1 = results %>%
  mutate(fitted = ifelse(median_P_value   > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x = results$outliersPro , y= results$median_P_value)) + 
  geom_point(aes(color = fitted , alpha = .5) ) +
    labs(title = "Outliers Proportion And P_value After"
                             )  

p2 = results %>%
  mutate(fitted = ifelse(median_P_value > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x= fitted)) + 
  geom_bar(aes(fill=fitted)) +
    labs(title = "Outliers Proportion And P_value After"
                             )

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )


ddd <-results %>%
  mutate(fittedBefore = ifelse(Before_P_value > 0.05,TRUE,FALSE)) %>%
  mutate(fittedAfter = ifelse(median_P_value > 0.05,TRUE,FALSE)) 

 
library(MASS)
tb1<-table(ddd$fittedBefore, ddd$fittedAfter)
tb1
chisq.test(tb1)


```




# Relation Between Outliers Proportion and P-value After {Q_b_F_C_P} 
```{r}

p1 = results %>%
  mutate(fitted = ifelse(Q_b_F_C_P_value   > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x = results$outliersPro , y= results$Q_b_F_C_P_value)) + 
  geom_point(aes(color = fitted , alpha = .5) ) +
    labs(title = "Outliers Proportion And P_value After"
                             )  

p2 = results %>%
  mutate(fitted = ifelse(Q_b_F_C_P_value > 0.05,TRUE,FALSE)) %>% 
  ggplot(aes(x= fitted)) + 
  geom_bar(aes(fill=fitted)) +
    labs(title = "Outliers Proportion And P_value After"
                             )

gridExtra::grid.arrange(p1,p2 ,  ncol = 2 )


ddd <-results %>%
  mutate(fittedBefore = ifelse(Before_P_value > 0.05,TRUE,FALSE)) %>%
  mutate(fittedAfter = ifelse(Q_b_F_C_P_value > 0.05,TRUE,FALSE)) 

 

library(MASS)
tb1<-table(ddd$fittedBefore, ddd$fittedAfter)
tb1
chisq.test(tb1)



```


$\frac{1}{n} \sum_{i=1}^n (\beta - \hat{\beta})^2$

###  Summations

### Without Indices

$\sum x_{i}$

$\sum x_{i}^2$

$\sum x_{i}y_{i}$

#### With Indices - Inline Form

$\sum_{i=1}^n x_{i}$

$\sum_{i=1}^n x_{i}^2$

$\sum_{i=1}^n x_{i}y_{i}$

#### With Indices - Display Form

$$\sum_{i=1}^n x_{i}y_{i}$$

### Independent Samples

$$\mu_{\bar{x_{1}} - \bar{x_{2}}} = \mu_{1} - \mu_{2}$$

$$\sigma_{\bar{x_{1}} - \bar{x_{2}}}^2 = \frac {\sigma_{1}^2}{n_{1}} + \frac{\sigma_{2}^2}{n_{2}}$$

$$\mu_{\hat{p}_{1} - \hat{p}_{2}} = p_{1} - p_{2}$$

$$\sigma_{\hat{p}_{1} - \hat{p}_{2}}^2 = \frac {p_{1}(1 - p_{1})}{n_{1}} + \frac {p_{2}(1 - p_{2})}{n_{2}}$$


### Pooled Sample Variance

$$s_{p}^2 = \frac {(n_{1} - 1)s_{1}^2 + (n_{2} - 1)s_{2}^2}{n_{1} + n_{2} - 2}$$

### Pooled Sample Proportion

$$\hat{p} = \frac {n_{1}\hat{p}_1 + n_{2}\hat{p}_{2}}{n_{1} + n_{2}}$$

### Chi-Square Test

$$\chi^2 = \sum \frac {(O - E)^2}{E}$$


### Correlations

$${SS}_{xx} = \sum (x - \bar{x})^2 = \sum x^2 - \frac {(\sum x)^2}{n}$$

$${SS}_{xy} = \sum (x - \bar{x})(y - \bar{y}) = \sum xy - \frac {(\sum x)(\sum y)}{n}$$

$$r = \frac {{SS}_{xy}}{\sqrt {{SS}_{xx}{SS}_{yy}}}$$


### Regression

#### Population Regression Line

$$E(y) = \alpha + \beta{x}$$

$$var(y) = \sigma^2$$

#### Least Squares Line

$$\hat{y} = a + bx$$

where 

$$b = \frac {{SS}_{xy}}{{SS}_{xx}}$$

and 

$$\bar{y} = a + b\bar{x}$$


#### Residual Sum of Squares

$$SSResid = \sum (y - \hat{y})^2 = \sum y^2 - a\sum y - b \sum xy$$

#### Standard Errors

$$s_{e} = \sqrt \frac {SSResid}{n - 2}$$

$$s_{b} = \frac {s_{e}}{\sqrt {{SS}_{xx}}}$$

$$s_{a + bx} = s_{e} \sqrt {1 + \frac {1}{n} + \frac {(x - \bar{x})^2}{{SS}_{xx}}}$$

for prediction:

$$se(y - \hat{y}) = s_{e} \sqrt {1 + \frac {1}{n} + \frac {(x - \bar{x})^2}{{SS}_{xx}}}$$


### Variance

$$SSTr = \frac {T_{1}^2}{n_{1}} + \frac {T_{2}^2}{n_{2}} + ... + \frac {T_{k}^2}{n_{k}} - \frac {T^2}{n}$$

$$SSTo = x_{1}^2 + x_{2}^2 + ... + x_{k}^2 - \frac {T^2}{n}$$

$$SSE = SSTo - SSTr$$


 


